---
title: "nano_string_ncounter_process"
author: "Cheng Sheng Lai"
date: "2024-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NanoString nCounter data analysis pipeline:
This is a R Markdown for processing gene expression data of NanoString nCounter platform.
The analysis pipeline is consist of 7 steps:
 (1) packages and function import
 (2) data import 
 (3) data preprocessing
 (4) data quality check
 (5) normalisation
 (6) downstream data analysis
 (7) visualisation

###To install all of the packages, this Markdown need to work in R version 4.0.0. 
###The pipeline requires the users to set up directory for data import and file storage.
#1. Packages and function import:
```{r}
#Import self-defined function required in the pipeline:
source('nanostring_RUVSeq_functions.R')

if (!require("BiocManager", quietly = TRUE))
install.packages("BiocManager")
packageVersion("BiocManager")

#Check the current version of R:
R.version.string

install.packages("locfit")
BiocManager::install("DESeq2",force = TRUE)
BiocManager::install("DESeq",force = TRUE)
BiocManager::install("RUVSeq",force = TRUE)
BiocManager::install("NanoStringDiff",force = TRUE)
BiocManager::install("nanostringr",force = TRUE)
BiocManager::install("NanoStringQCPro",force = TRUE)
BiocManager::install("NanoStringNorm",force = TRUE)
require("locfit")
require("DESeq")
require("DESeq2")
require("RUVSeq")
require("nanostringr")
require("NanoStringDiff")
require("png")
require("NanoStringQCPro")
require("ggplot2")
require("EnvStats")
require("NanoStringNorm")
require("MASS")
require("dplyr")
require("formattable")
require("xfun")
require("limma")
require("matrixStats")
require("tidyr")
require("tibble")
require("Biobase")
require("GEOquery")
require("RUVSeq")
require("DESeq2")
require("matrixStats")
require("NanoStringNorm")
require("patchwork")
require("limma")
require("statmod")
require("VennDiagram")
require("gridExtra")
require("pheatmap")
require("viridis")
```

#2. Data import:
```{r}
##Create a list to document all the files name in in all of sub-directory:
#Set up the main directory with multiple folders contained RCC files:
setwd("RCC_files_folder")
main_directory <- "RCC_files_folder"

sub_directories <- list.dirs(path = main_directory, full.names = TRUE, recursive = FALSE)

sub_directories

all_rcc_files <- list()

for (sub_dir in sub_directories) {
  files <- list.files(path = sub_dir, full.names = TRUE)
  all_rcc_files <- c(all_rcc_files, files)
}

all_rcc_files

all_rcc_files = all_rcc_files[grepl('RCC',files.RCC)]

all_rcc_files <- unlist(all_rcc_files)

setwd(dirname(all_rcc_files[1]))

ng_1 = nrow(readRcc(basename(all_rcc_files[1]))$Code_Summary)

ncol_1 = length(all_rcc_files)


raw_expression_1 = as.data.frame(matrix(nrow = ng_1,ncol = length(all_rcc_files)+2))

colnames(raw_expression_1)[1:2] = c('Gene','Class')

pData_1 = as.data.frame(matrix(nrow = length(all_rcc_files),ncol = 15))

colnames(pData_1) = c('File_name','SampleID','Owner','Comments','Date','GeneRLF','SystemAPF','imagingQC','imagingQC_value', 'bindingDensityQC', 'bindingDensityQC_value', 'limitOfDetectionQC', 'limitOfDetectionQC_value', 'positiveLinearityQC', 'positiveLinearityQC_value')

raw_expression_1[,1:2] = readRcc(basename(all_rcc_files[1]))$Code_Summary[,c(2,1)]

```

#3. Data preprocessing and quality check:
```{r}

for (i in 1:length(all_rcc_files)){
  
  print(i)
  
  setwd(dirname(all_rcc_files[i]))
  
  rcc = readRcc(basename(all_rcc_files[i]))
  
  raw = rcc$Code_Summary
  
  raw_expression_1[,i+2] = as.numeric(raw$Count)
  
  colnames(raw_expression_1)[i+2] = strsplit(basename(all_rcc_files[i]), '_')[[1]][1]
  
  pData_1[i,2:7] = as.vector(rcc$Sample_Attributes)
  
  pData_1$imagingQC[i] = imagingQC(rcc)
  pData_1$imagingQC_value[i] = imagingQC_value(rcc)
  
  pData_1$bindingDensityQC[i] = bindingDensityQC(rcc,.1,1.8)
  pData_1$bindingDensityQC_value[i] = bindingDensityQC_value(rcc)
  
  pData_1$limitOfDetectionQC[i] = limitOfDetectionQC(rcc)
  pData_1$limitOfDetectionQC_value[i] = limitOfDetectionQC_value(rcc)

  pData_1$positiveLinearityQC[i] = positiveLinQC(rcc)
  pData_1$positiveLinearityQC_value[i] = positiveLinQC_value(rcc)
  
}


colnames(raw_expression_1) <- make.unique(colnames(raw_expression_1))

colname_raw_exp_1 <- colnames(raw_expression_1)[3:ncol(raw_expression_1)]

pData_1$File_name <- colname_raw_exp_1 


matches <- grep("^CHIVAS #\\d{1}$", pData_1$SampleID)
pData_1$SampleID[matches] <- gsub("^CHIVAS #(\\d{1})$", "CHIVAS #0\\1", pData_1$SampleID[matches])

matches <- grep("^CHIVAS #\\d{2}$", pData_1$SampleID)
pData_1$SampleID[matches] <- gsub("^CHIVAS #(\\d{2})$", "CHIVAS #0\\1", pData_1$SampleID[matches])

matches <- grep("^CHIVAS #\\d{3}$", pData_1$SampleID)
pData_1$SampleID[matches] <- gsub("^CHIVAS #(\\d{3})$", "CHIVAS-\\1", pData_1$SampleID[matches])

matches <- grep("^CHIVAS-\\d{3}$", pData_1$SampleID)
pData_1$SampleID[matches] <- gsub("^CHIVAS-(\\d{3})$", "\\1", pData_1$SampleID[matches])

Sample <- c("061", "063", "098", "100", "107", "109", "116", "117", "123", "127", "133", "139")

pData_1 <- pData_1 %>%
  mutate(SampleID = case_when(
    SampleID == "Sample01" ~ Sample[1],
    SampleID == "Sample02" ~ Sample[2],
    SampleID == "Sample03" ~ Sample[3],
    SampleID == "Sample04" ~ Sample[4],
    SampleID == "Sample05" ~ Sample[5],
    SampleID == "Sample06" ~ Sample[6],
    SampleID == "Sample07" ~ Sample[7],
    SampleID == "Sample08" ~ Sample[8],
    SampleID == "Sample09" ~ Sample[9],
    SampleID == "Sample10" ~ Sample[10],
    SampleID == "Sample11" ~ Sample[11],
    SampleID == "Sample12" ~ Sample[12],
    TRUE ~ SampleID 
  ))

pData_1 <- pData_1 %>% arrange(SampleID)

#Import "phenotype" csv file:
phenotype <- read.csv("phenotype.csv")

matches <- grep("^[0-9]$", phenotype$SampleID)
phenotype$SampleID[matches] <- gsub("(^[0-9])", "0\\1", phenotype$SampleID[matches])
matches <- grep("^[0-9]{2}$", phenotype$SampleID)
phenotype$SampleID[matches] <- gsub("(^[0-9]{2}$)", "0\\1", phenotype$SampleID[matches])

phenotype <- phenotype%>%arrange(SampleID)

phenotype_filter <- semi_join(phenotype, pData_1, by = "SampleID")

colnames(phenotype_filter)
phenotype_selected <- phenotype_filter %>% dplyr::select(SampleID, Experiment:Time_of_diagnosis_or_peak)

merged_pData_1 <- inner_join(phenotype_selected, pData_1, by = "SampleID")

cols_pData_1 <- setdiff(names(pData_1), "SampleID")
cols_phenotype_selected <- setdiff(names(phenotype_selected), "SampleID")
new_merged_pData_1 <- c("SampleID", cols_pData_1, cols_phenotype_selected)
merged_pData_1 <- merged_pData_1 %>% dplyr::select(all_of(new_merged_pData_1))

merged_pData_1 <- merged_pData_1 %>% filter((SampleID != "PanelStandard"))

#Deplete the flagged sample:
merged_pData_1 <- merged_pData_1 %>% filter(!(imagingQC_value %in% "NaN"|bindingDensityQC_value %in% "NaN"|limitOfDetectionQC_value %in% "NaN"|positiveLinearityQC_value %in% "NaN"))

duplicate_sample <- merged_pData_1 %>%
  group_by(SampleID) %>%
  filter(n() > 1) %>%
  ungroup()

#Remove the duplicates that with: 
#(1) bindingDensityQC_value below 0.1 
#(2) higher value of "Limit of detection":
max_lod_sample <- duplicate_sample %>%
  group_by(SampleID) %>%
  arrange(SampleID, bindingDensityQC_value >= 0.1, desc(limitOfDetectionQC_value)) %>%
  slice(1) %>%
  ungroup()

merged_pData_1 <- anti_join(merged_pData_1, max_lod_sample, by = c("SampleID", "limitOfDetectionQC_value"))
```

#4. Data quality check:
```{r}
###Decide which data point should be removed based on the result of QC. 
#Label the flag value in the dataframe "pData_1":
formattable(merged_pData_1, list(
  imagingQC = formatter("span",
                        style = x ~ ifelse(x == "Flag", 
                                           style(color = "red", font.weight = "bold"), 
                                           style(color = "green"))),
  imagingQC_value = formatter("span",
                              style = x ~ ifelse(is.na(x), NA,
                                                 ifelse(merged_pData_1$imagingQC == "Flag", 
                                                        style(color = "red", font.weight = "bold"), 
                                                        style(color = "green")))),
  bindingDensityQC = formatter("span",
                               style = x ~ ifelse(x == "Flag", 
                                                  style(color = "red", font.weight = "bold"), 
                                                  style(color = "green"))),
  bindingDensityQC_value = formatter("span",
                                     style = x ~ ifelse(is.na(x), NA,
                                                        ifelse(merged_pData_1$bindingDensityQC == "Flag", 
                                                               style(color = "red", font.weight = "bold"), 
                                                               style(color = "green")))),
  limitOfDetectionQC = formatter("span",
                                 style = x ~ ifelse(x == "Flag",
                                                    style(color = "red", font.weight = "bold"), 
                                                    style(color = "green"))),
  limitOfDetectionQC_value = formatter("span",
                                       style = x ~ ifelse(is.na(x), NA,
                                                          ifelse(merged_pData_1$limitOfDetectionQC == "Flag", 
                                                                 style(color = "red", font.weight = "bold"), 
                                                                 style(color = "green")))),
  positiveLinearityQC = formatter("span",
                                  style = x ~ ifelse(x == "Flag",
                                                     style(color = "red", font.weight = "bold"), 
                                                     style(color = "green"))),
  positiveLinearityQC_value = formatter("span",
                                        style = x ~ ifelse(is.na(x), NA,
                                                           ifelse(merged_pData_1$positiveLinearityQC == "Flag", 
                                                                  style(color = "red", font.weight = "bold"), 
                                                                  style(color = "green")))),
  SampleID = formatter("span",
                       style = x ~ ifelse(merged_pData_1$imagingQC == "Flag"|merged_pData_1$bindingDensityQC == "Flag"|merged_pData_1$limitOfDetectionQC == "Flag"|merged_pData_1$positiveLinearityQC == "Flag",
                                          style(color = "red", font.weight = "bold"),
                                          NA))
))

#Trim the dataframe "raw_expression_1" with the column "File_name" in "merged_pData_1":
selected_columns <- c("Gene", "Class", merged_pData_1$File_name)
raw_expression_1 <- raw_expression_1 %>% select(all_of(selected_columns))

```
## Here, the dataframes "raw_expression_1" and "merged_pData_1" are completed.
##"raw_expression_1": Gene expression counts of each file.
##"merged_pData_1": Detail information for each file.

#5. Normalisation_1:
```{r}
raw_1 = raw_expression_1[,-c(1:2)]
fData_1 = raw_expression_1[,c(1:2)]

#Collect all the genes that be classified as "Housekeeping" genes, and defined as "cIdx":
##Deplete the Housekeeping gene below 100:
cIdx <- fData_1$Gene[fData_1$Class == "Housekeeping"]

merged_pData_1$HK_Gene_Miss = colSums(raw_1[cIdx,] == 0)

rownames(fData_1) = fData_1$Gene

rownames(raw_1) = fData_1$Gene

rownames(merged_pData_1) = colnames(raw_1)

####Check if housekeeping genes are associated with primary phenotypes:
hk_raw_1 = raw_1[cIdx,]

pval_1 = vector(length = nrow(hk_raw_1))

for (i in 1:nrow(hk_raw_1)){
  
  reg = glm.nb(as.numeric(hk_raw_1[i,]) ~ as.factor(merged_pData_1$Outcome4))
  
  pval_1[i] = coef(summary(reg))[2,4]
}

sum(pval_1 <= .05)
pval_1

##Based on the the result of negative binomial GLM model test to remove housekeeping genes.

raw_log <- data.frame(raw_1[1:773,])

rld_data_raw <- log2(raw_log)

gene_medians_raw <- apply(rld_data_raw, 1, median)

rle_data_raw <- sweep(rld_data_raw, 1, gene_medians_raw)

colnames(rle_data_raw) <- make.unique(colnames(rle_data_raw))
rle_data_raw_df <- as.data.frame(rle_data_raw)
rle_data_raw_df <- rownames_to_column(rle_data_raw_df, var = "gene")
rle_long_raw <- pivot_longer(rle_data_raw_df, cols = -gene, names_to = "sample", values_to = "value")

##Normalize data with "RUVseq" method:
k = 1
vsd = RUV_total(raw_1,merged_pData_1,fData_1,k = k)$vsd
set = RUV_total(raw_1,merged_pData_1,fData_1,k = k)$set 
save(vsd,file = paste0("strep_A_ruv_vsd_k_1", k,".rda"))
save(set,file = paste0("strep_A_ruv_set_k_1", k,".rda"))


i = 1
results = paste0("strep_A_deg_k_1",i,".csv")
load(paste0("strep_A_ruv_set_k_1",i,".rda"))

dds_1 <- DESeqDataSetFromMatrix(countData = counts(set)[1:773,],
                              colData = pData(set),
                              design = ~ 1)

dds_1 <- DESeq(dds_1)

size_factors_1 <- sizeFactors(dds_1)
print(size_factors_1)

rld_1_normalized <- counts(dds_1, normalized=TRUE)
rld_1_normalized <- log2(rld_1_normalized)
rld_1_normalized[is.infinite(rld_1_normalized) & rld_1_normalized < 0] <- 0
rld_1_normalized[is.na(rld_1_normalized)] <- 0

gene_medians_1 <- apply(rld_1_normalized, 1, median)

rle_data_1 <- sweep(rld_1_normalized, 1, gene_medians_1)

colnames(rle_data_1) <- make.unique(colnames(rle_data_1))
rle_data_1_df <- as.data.frame(rle_data_1)
rle_data_1_df <- rownames_to_column(rle_data_1_df, var = "gene")
rle_long_1 <- pivot_longer(rle_data_1_df, cols = -gene, names_to = "sample", values_to = "value")

##Create a Dataset for DESeq analysis (nSolver):
dds.nsolver_1 <- DESeqDataSetFromMatrix(countData = counts(set)[1:773,],
                                        colData = pData(set),
                                        design = ~1)
pos_1 = raw_1[grepl('Positive',raw_expression_1$Class),]
hk_1 = raw_1[grepl('Housekeeping',raw_expression_1$Class),]
pos.factors_1 = mean(as.numeric(apply(pos_1,2,geoMean)))/as.numeric(apply(pos_1,2,geoMean))
hk.factors_1 = mean(as.numeric(apply(hk_1,2,geoMean)))/as.numeric(apply(hk_1,2,geoMean))
sizeFactors(dds.nsolver_1) <- pos.factors_1 * hk.factors_1
dds.nsolver_1 <- DESeq(dds.nsolver_1)

size_factors_2 <- sizeFactors(dds.nsolver_1)
print(size_factors_2)

rld_2_normalized <- counts(dds.nsolver_1, normalized=TRUE)
rld_2_normalized <- log2(rld_2_normalized)
rld_2_normalized[is.infinite(rld_2_normalized) & rld_2_normalized < 0] <- 0
rld_2_normalized[is.na(rld_2_normalized)] <- 0

gene_medians_2 <- apply(rld_2_normalized, 1, median)

rle_data_2 <- sweep(rld_2_normalized, 1, gene_medians_2)

colnames(rle_data_2) <- make.unique(colnames(rle_data_2))
rle_data_2_df <- as.data.frame(rle_data_2)
rle_data_2_df <- rownames_to_column(rle_data_2_df, var = "gene")
rle_long_2 <- pivot_longer(rle_data_2_df, cols = -gene, names_to = "sample", values_to = "value")

##Check whether the normalization result from nSolver software aligns with the algorithm: 
nSolver_result <- read.csv("nSolver_analysed_result") 
rownames(nSolver_result) <- nSolver_result[,1]
nSolver_result <- nSolver_result[,-1]
nSolver_result <- log2(nSolver_result)

gene_medians_3 <- apply(nSolver_result, 1, median)

rle_data_3 <- sweep(nSolver_result, 1, gene_medians_2)

colnames(rle_data_3) <- make.unique(colnames(rle_data_3))
rle_data_3_df <- as.data.frame(rle_data_3)
rle_data_3_df$gene <- rownames(rle_data_3_df) 
rle_long_3 <- pivot_longer(rle_data_3_df, cols = -gene, names_to = "sample", values_to = "value")

```

#5. Normalisation_2(RLE plot):
```{r}
#Raw data RLE plot:
y_limits <- c(-8, 8)
y_breaks <- seq(-8,8, by = 2)

rle_plot_raw <- ggplot(rle_long_raw, aes(x = sample, y = value)) +
              geom_boxplot(coef = 6, outlier.shape = 16, color = "black") +
              scale_y_continuous(limits = y_limits, breaks = y_breaks) + 
              stat_summary(fun = median, geom = "line", size = 4) +
              geom_hline(yintercept = 0, color = "red", size = .5) +
              theme_minimal() +
              labs(title = "RLE Plot (Raw)",
                   x = NULL,
                   y = NULL) +
              theme(axis.text.x = element_blank(),
                    axis.text.y = element_text(size = 18),
                    axis.ticks.x = element_blank(),
                    title = element_text(size = 18, face = "bold"))

rle_plot_raw

#RUVseq normalised data RLE plot: 
rle_plot_1 <- ggplot(rle_long_1, aes(x = sample, y = value)) +
               geom_boxplot(coef = 6, outlier.shape = 16, color = "black") +
               scale_y_continuous(limits = y_limits, breaks = y_breaks) + 
               stat_summary(fun = median, geom = "line", size = 4) +
               geom_hline(yintercept = 0, color = "red", size = .5) +
               theme_minimal() +
               labs(title = "RLE Plot (RUVseq)",
                    x = NULL,
                    y = "Median deviation of log2 expression") +
               theme(axis.text.x = element_blank(),
                     axis.text.y = element_text(size = 18),
                     axis.ticks.x = element_blank(),
                     axis.title.y = element_text(size = 18),
                     title = element_text(size = 18, face = "bold"))

rle_plot_1

#nSolver normalised data RLE plot:
rle_plot_2 <- ggplot(rle_long_2, aes(x = sample, y = value)) +
                     geom_boxplot(coef = 6, outlier.shape = 16, color = "black") +
                     scale_y_continuous(limits = y_limits, breaks = y_breaks) + 
                     stat_summary(fun = median, geom = "line", size = 4) +
                     geom_hline(yintercept = 0, color = "red", size = .5) +
                     theme_minimal() +
                     labs(title = "RLE Plot (nSolver)",
                          x = "Sample",
                          y = NULL) +
                     theme(axis.text.x = element_blank(),
                           axis.text.y = element_text(size = 14), 
                           axis.ticks.x = element_blank(),
                           axis.title.x = element_text(size = 16),
                           title = element_text(size = 18, face = "bold"))

rle_plot_2

#nSolver software normalised data RLE plot:
rle_plot_3 <- ggplot(rle_long_3, aes(x = sample, y = value)) +
  geom_boxplot(coef = 6, outlier.shape = 16, color = "black") +
  scale_y_continuous(limits = y_limits, breaks = y_breaks) + 
  stat_summary(fun = median, geom = "line", size = 4) +
  geom_hline(yintercept = 0, color = "red", size = .5) +
  theme_minimal() +
  labs(title = "RLE Plot (nSolver software)",
       x = "Sample",
       y = NULL) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 18),
        axis.ticks.x = element_blank(),
        axis.title.x = element_text(size = 22, face = "bold"),
        title = element_text(size = 18, face = "bold"))

rle_plot_3

##Merge "rle_plot_raw", "rle_plot_1" and "rle_plot_2", "rle_plot_3" into one figure:
RLE_plot_rlog <- rle_plot_raw/rle_plot_1/rle_plot_3
setwd("RCC_files_folder")
ggsave("RLE_plot.pdf", plot = RLE_plot_rlog, device = "pdf", width = 18, height = 12)
RLE_plot_rlog

```

#6. Downstream analysis:
```{r}
##Create analysis model in "Timepoint" and "Symptoms" with limma: 
raw_dataframe <- rld_1_normalized
pheno_dataframe <- AnnotatedDataFrame(data.frame(colData(dds_1)))
feature_dataframe <- AnnotatedDataFrame(data.frame(fData_1[1:773, ]))

dds_limma <- ExpressionSet(assayData = raw_dataframe,
                           phenoData = pheno_dataframe,
                           featureData = feature_dataframe)

dim(dds_limma)

dds_limma$Timepoint <- as.character(dds_limma$Timepoint)

exprs_data <- exprs(dds_limma)
pData_RUV <- pData(dds_limma)
plotDensities(exprs_data, legend = FALSE)


group_time_outcome_RUV <- with(pData(dds_limma), paste(Outcome2, Timepoint, sep = "."))
group_time_outcome_RUV <- gsub(" ", "_", group_time_outcome_RUV)
group_time_outcome_RUV <- factor(group_time_outcome_RUV)
design_RUV <- model.matrix(~0 + group_time_outcome_RUV, data = pData(dds_limma))

corfit_participant <- duplicateCorrelation(exprs_data, design_RUV, block = pData_RUV$Participant)

colnames(design_RUV) <- levels(group_time_outcome_RUV)
colSums(design_RUV)
cm <- makeContrasts(
  #Same outcome with different timepoint:
  T24_vs_T0_NP = No_Pharyngitis.24 - No_Pharyngitis.0,
  T36_vs_T0_NP = No_Pharyngitis.36 - No_Pharyngitis.0,
  T48_vs_T0_NP = No_Pharyngitis.48 - No_Pharyngitis.0,
  T60_vs_T0_NP = No_Pharyngitis.60 - No_Pharyngitis.0,
  T72_vs_T0_NP = No_Pharyngitis.72 - No_Pharyngitis.0,
  T84_vs_T0_NP = No_Pharyngitis.84 - No_Pharyngitis.0,
  T96_vs_T0_NP = No_Pharyngitis.96 - No_Pharyngitis.0,
  T108_vs_T0_NP = No_Pharyngitis.108 - No_Pharyngitis.0,
  
  T24_vs_T0_P = Pharyngitis.24 - Pharyngitis.0,
  T36_vs_T0_P = Pharyngitis.36 - Pharyngitis.0,
  T48_vs_T0_P = Pharyngitis.48 - Pharyngitis.0,
  T60_vs_T0_P = Pharyngitis.60 - Pharyngitis.0,
  T72_vs_T0_P = Pharyngitis.72 - Pharyngitis.0,
  T84_vs_T0_P = Pharyngitis.84 - Pharyngitis.0,
  T96_vs_T0_P = Pharyngitis.96 - Pharyngitis.0,
  T120_vs_T0_P = Pharyngitis.120 - Pharyngitis.0,
  T132_vs_T0_P = Pharyngitis.132 - Pharyngitis.0,
  #Same timepoint with different outcome:
  P_vs_NP_T0 = Pharyngitis.0 - No_Pharyngitis.0,
  P_vs_NP_T24 = Pharyngitis.24 - No_Pharyngitis.24,
  P_vs_NP_T48 = Pharyngitis.48 - No_Pharyngitis.48,
  P_vs_NP_T60 = Pharyngitis.60 - No_Pharyngitis.60,
  P_vs_NP_T72 = Pharyngitis.72 - No_Pharyngitis.72,
  P_vs_NP_T84 = Pharyngitis.84 - No_Pharyngitis.84,
  P_vs_NP_T96 = Pharyngitis.96 - No_Pharyngitis.96,
  P_vs_NP_T120 = Pharyngitis.120 - No_Pharyngitis.120,
  #Different timepoint with differnent outcome:
  T0_P_vs_T0_NP = Pharyngitis.0 - No_Pharyngitis.0,
  T24_P_vs_T0_NP = Pharyngitis.24 - No_Pharyngitis.0,
  T36_P_vs_T0_NP = Pharyngitis.36 - No_Pharyngitis.0,
  T48_P_vs_T0_NP = Pharyngitis.48 - No_Pharyngitis.0,
  T60_P_vs_T0_NP = Pharyngitis.60 - No_Pharyngitis.0,
  T72_P_vs_T0_NP = Pharyngitis.72 - No_Pharyngitis.0,
  T84_P_vs_T0_NP = Pharyngitis.84 - No_Pharyngitis.0,
  T96_P_vs_T0_NP = Pharyngitis.96 - No_Pharyngitis.0,
  T120_P_vs_T0_NP = Pharyngitis.120 - No_Pharyngitis.0,
  T132_P_vs_T0_NP = Pharyngitis.132 - No_Pharyngitis.0,
  levels = design_RUV)

fit_RUV <- lmFit(dds_limma, design_RUV, block = pData_RUV$Participant, correlation = corfit_participant$consensus.correlation)
fit_RUV_1 <- contrasts.fit(fit_RUV, contrasts = cm)
fit_RUV_1 <- eBayes(fit_RUV_1)
results_RUV <- decideTests(fit_RUV_1)
summary(results_RUV)

#Create volcanoplot to display gene expression differentiation:
topTable_results_P_vs_NP_T0 <- topTable(fit_RUV_1, coef="P_vs_NP_T0", number=Inf)
topTable_results_P_vs_NP_T24 <- topTable(fit_RUV_1, coef="P_vs_NP_T24", number=Inf)
topTable_results_P_vs_NP_T48 <- topTable(fit_RUV_1, coef="P_vs_NP_T48", number=Inf)
topTable_results_P_vs_NP_T60 <- topTable(fit_RUV_1, coef="P_vs_NP_T60", number=Inf)
topTable_results_P_vs_NP_T72 <- topTable(fit_RUV_1, coef="P_vs_NP_T72", number=Inf)
topTable_results_P_vs_NP_T120 <- topTable(fit_RUV_1, coef="P_vs_NP_T120", number=Inf)

```

#7. Visualisation_1(Venn diagram):
```{r}
##Demonstrate the number of up- and down-regulation genes in each comparison groups:
#Create up- and down-regulation list for "topTable_results_P_vs_NP_T24":
T24_up <- topTable_results_P_vs_NP_T24 %>% 
  filter(logFC > 2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T24_down <- topTable_results_P_vs_NP_T24 %>% 
  filter(logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)

T24_up <- as.character(T24_up$Gene)
T24_down <- as.character(T24_down$Gene)
#Create up- and down-regulation list for "topTable_results_P_vs_NP_T48":
T48_up <- topTable_results_P_vs_NP_T48 %>% 
  filter(logFC > 2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T48_down <- topTable_results_P_vs_NP_T48 %>% 
  filter(logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)

T48_up <- as.character(T48_up$Gene)
T48_down <- as.character(T48_down$Gene)
#Create up- and down-regulation list for "topTable_results_P_vs_NP_T60":
T60_up <- topTable_results_P_vs_NP_T60 %>% 
  filter(logFC > 2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T60_down <- topTable_results_P_vs_NP_T60 %>% 
  filter(logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)

T60_up <- as.character(T60_up$Gene)
T60_down <- as.character(T60_down$Gene)
#Create up- and down-regulation list for "topTable_results_P_vs_NP_T72":
T72_up <- topTable_results_P_vs_NP_T72 %>% 
  filter(logFC > 2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T72_down <- topTable_results_P_vs_NP_T72 %>% 
  filter(logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)

T72_up <- as.character(T72_up$Gene)
T72_down <- as.character(T72_down$Gene)
#Create up- and down-regulation list for "topTable_results_P_vs_NP_T120":
T120_up <- topTable_results_P_vs_NP_T120 %>% 
  filter(logFC > 2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T120_down <- topTable_results_P_vs_NP_T120 %>% 
  filter(logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)

T120_up <- as.character(T120_up$Gene)
T120_down <- as.character(T120_down$Gene)

setwd("RCC_files_folder")

venn.plot_up <- venn.diagram(
  x = list(
    "24 hr" = T24_up, 
    "48 hr" = T48_up, 
    "60 hr" = T60_up, 
    "72 hr" = T72_up, 
    "120 hr" = T120_up
  ),
  category.names = c("24 hr", "48 hr", "60 hr", "72 hr", "120 hr"),
  filename = NULL,
  fill = c("red", "orange", "yellow", "green", "blue"),
  alpha = 0.5,
  cex = 2,
  fontfamily = "serif",
  fontface = "bold",
  cat.cex = 3,
  cat.fontfamily = "serif",
  cat.fontface = "bold",
  cat.dist = rep(0.05, 5), 
  cat.pos = c(-1, 1, -1, 1, -1) 
)

up_values_count <- c(
  "24 hr" = length(T24_up),
  "48 hr" = length(T48_up),
  "60 hr" = length(T60_up),
  "72 hr" = length(T72_up),
  "120 hr" = length(T120_up)
)

up_values_count <- data.frame(
  Group = names(up_values_count),
  Gene = up_values_count,
  stringsAsFactors = FALSE
)

up_table_grob <- tableGrob(up_values_count, rows = NULL, theme = ttheme_default(base_size = 14))

up_combined_plot <- arrangeGrob(gTree(children = venn.plot_up), up_table_grob,
                                ncol = 2, widths = c(4, 1),
                                top = textGrob("Upregulated gene",
                                               gp = gpar(fontsize = 20, fontface = "bold")))

grid.newpage()
grid.draw(up_combined_plot)
setwd("RCC_files_folder")
ggsave("up_venn_diagram.pdf", plot = up_combined_plot, device = "pdf", width = 15, height = 10)

venn.plot_down <- venn.diagram(
  x = list(
    "24 hr" = T24_down, 
    "48 hr" = T48_down, 
    "60 hr" = T60_down, 
    "72 hr" = T72_down, 
    "120 hr" = T120_down
  ),
  category.names = c("24 hr", "48 hr", "60 hr", "72 hr", "120 hr"),
  filename = NULL,
  fill = c("red", "orange", "yellow", "green", "blue"),
  alpha = 0.5,
  cex = 2,
  fontfamily = "serif",
  fontface = "bold",
  cat.cex = 3,
  cat.fontfamily = "serif",
  cat.fontface = "bold",
  cat.dist = rep(0.05, 5), 
  cat.pos = c(-1, 1, -1, 1, -1) 
)

down_values_count <- c(
  "24 hr" = length(T24_down),
  "48 hr" = length(T48_down),
  "60 hr" = length(T60_down),
  "72 hr" = length(T72_down),
  "120 hr" = length(T120_down)
)

down_values_count <- data.frame(
  "Group" = names(down_values_count),
  "Gene" = down_values_count,
  stringsAsFactors = FALSE
)

down_table_grob <- tableGrob(down_values_count, rows = NULL, theme = ttheme_default(base_size = 14))

down_combined_plot <- arrangeGrob(gTree(children = venn.plot_down), down_table_grob,
                                ncol = 2, widths = c(4, 1),
                                top = textGrob("Downregulated gene",
                                               gp = gpar(fontsize = 20, fontface = "bold")))


grid.newpage()
grid.draw(down_combined_plot)
ggsave("down_venn_diagram.pdf", plot = down_combined_plot, device = "pdf", width = 15, height = 10)

```

#7. Visualisation_2(Volcano plot):
```{r}
##Create up-regulation list of all timepoint:
T24_up_df <- data.frame(Value = T24_up, Gene = T24_up, stringsAsFactors = FALSE)
T48_up_df <- data.frame(Value = T48_up, Gene = T48_up, stringsAsFactors = FALSE)
T60_up_df <- data.frame(Value = T60_up, Gene = T60_up, stringsAsFactors = FALSE)
T72_up_df <- data.frame(Value = T72_up, Gene = T72_up, stringsAsFactors = FALSE)
T120_up_df <- data.frame(Value = T120_up, Gene = T120_up, stringsAsFactors = FALSE)

up_all <- unique(c(T24_up, T48_up, T60_up, T72_up, T120_up))
up_list <- data.frame(Gene = up_all, stringsAsFactors = FALSE)
up_list <- up_list %>%
  left_join(T24_up_df, by = "Gene") %>%
  left_join(T48_up_df, by = "Gene") %>%
  left_join(T60_up_df, by = "Gene") %>%
  left_join(T72_up_df, by = "Gene") %>%
  left_join(T120_up_df, by = "Gene")

rownames(up_list) <- up_list$Gene
up_list <- up_list[, -1]
colnames(up_list) <- c("T24", "T48", "T60", "T72", "T120")
up_non_na_counts <- apply(up_list, 1, function(row) sum(!is.na(row)))
up_list <- up_list[order(up_non_na_counts, decreasing = TRUE), ]


##Create down-regulation list of all timepoint:
T24_down_df <- data.frame(Value = T24_down, Gene = T24_down, stringsAsFactors = FALSE)
T48_down_df <- data.frame(Value = T48_down, Gene = T48_down, stringsAsFactors = FALSE)
T60_down_df <- data.frame(Value = T60_down, Gene = T60_down, stringsAsFactors = FALSE)
T72_down_df <- data.frame(Value = T72_down, Gene = T72_down, stringsAsFactors = FALSE)
T120_down_df <- data.frame(Value = T120_down, Gene = T120_down, stringsAsFactors = FALSE)

down_all <- unique(c(T24_down, T48_down, T60_down, T72_down, T120_down))
down_list <- data.frame(Gene = down_all, stringsAsFactors = FALSE)
down_list <- down_list %>%
  left_join(T24_down_df, by = "Gene") %>%
  left_join(T48_down_df, by = "Gene") %>%
  left_join(T60_down_df, by = "Gene") %>%
  left_join(T72_down_df, by = "Gene") %>%
  left_join(T120_down_df, by = "Gene")

rownames(down_list) <- down_list$Gene
down_list <- down_list[, -1]
colnames(down_list) <- c("T24", "T48", "T60", "T72", "T120")
down_non_na_counts <- apply(down_list, 1, function(row) sum(!is.na(row)))
down_list <- down_list[order(down_non_na_counts, decreasing = TRUE), ]

##Visualise the gene expression difference in each comparison groups with volcano plot:
gene_RUV <- fit_RUV_1$genes[, "Gene"]

T0_vol <- volcanoplot_neo(fit_RUV_1, coef = "P_vs_NP_T0", highlight = 10,
                          names = gene_RUV, title = "(a) 0 hrs gene expression comparison",
                          legend = FALSE)
T24_vol <- volcanoplot_neo(fit_RUV_1, coef = "P_vs_NP_T24", highlight = 10,
                           names = gene_RUV, title = "(b) 24 hrs gene expression comparison",
                           legend = FALSE)
T48_vol <- volcanoplot_neo(fit_RUV_1, coef = "P_vs_NP_T48", highlight = 10,
                           names = gene_RUV, title = "(c) 48 hrs gene expression comparison",
                           legend = FALSE)
T60_vol <- volcanoplot_neo(fit_RUV_1, coef = "P_vs_NP_T60", highlight = 10, 
                           names = gene_RUV, title = "(d) 60 hrs gene expression comparison",
                           legend = FALSE)
T72_vol <- volcanoplot_neo(fit_RUV_1, coef = "P_vs_NP_T72", highlight = 10,
                           names = gene_RUV, title = "(e) 72 hrs gene expression comparison",
                           legend = FALSE)
T120_vol <- volcanoplot_neo(fit_RUV_1, coef = "P_vs_NP_T120", highlight = 10,
                           names = gene_RUV, title = "(f) 120 hrs gene expression comparison",
                           legend = FALSE)

vol_final <- (T0_vol|T24_vol) / (T48_vol|T60_vol) / (T72_vol|T120_vol)
vol_final
ggsave("Vol_plot.pdf", plot = vol_final, device = "pdf", width = 21, height = 14)

```

#7. Visualisation_2(Heatmap):
```{r}
##Create heatamp to demonstrate gene expression among all groups: 

annotation_col <- merged_pData_1[, c("Timepoint", "Outcome2")]
names(annotation_col)[names(annotation_col) == "Outcome2"] <- "Symptom"

rownames(annotation_col) <- colnames(raw_dataframe)

annotation_col <- arrange(annotation_col, Symptom, Timepoint)

gene_means_1 <- apply(rld_1_normalized, 1, mean)
normalized_raw_dataframe <- sweep(rld_1_normalized, 1, gene_means_1)

arrange_raw_dataframe <- normalized_raw_dataframe[, rownames(annotation_col)]
arrange_raw_dataframe <- as.data.frame(arrange_raw_dataframe)

T0_sig <- topTable_results_P_vs_NP_T0 %>% 
  filter(logFC > 2 | logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T24_sig <- topTable_results_P_vs_NP_T24 %>% 
  filter(logFC > 2 | logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T48_sig <- topTable_results_P_vs_NP_T48 %>% 
  filter(logFC > 2 | logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T60_sig <- topTable_results_P_vs_NP_T60 %>% 
  filter(logFC > 2 | logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T72_sig <- topTable_results_P_vs_NP_T72 %>% 
  filter(logFC > 2 | logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)
T120_sig <- topTable_results_P_vs_NP_T120 %>% 
  filter(logFC > 2 | logFC < -2, P.Value < 0.05) %>% 
  arrange(desc(logFC)) %>%
  select(Gene)

sig_all <- bind_rows(T0_sig, T24_sig, T48_sig, T60_sig, T72_sig, T120_sig) %>%
           distinct()

arrange_raw_dataframe <- arrange_raw_dataframe %>%
                         filter(rownames(arrange_raw_dataframe) %in% rownames(sig_all))

arrange_raw_dataframe <- arrange_raw_dataframe %>%
                         arrange(match(rownames(arrange_raw_dataframe),rownames(sig_all)))


##Plot the heat map of gene expression results:

Heatmap_Strep_A <- pheatmap(arrange_raw_dataframe,
         main = "Gene expression of Streptococcus infection",
         cluster_cols = FALSE,
         cutree_rows = 3,
         gaps_col = c(6, 12, 13, 19, 22, 28, 29, 35, 36, 42, 61, 78, 84, 99, 101, 104, 105, 106, 121),
         show_colnames = F,
         show_rownames = T,
         col = viridis(10, option = "plasma"),
         border_color = "black",
         annotation_col = annotation_col)

ggsave("Heatmap_Strep_A.pdf", plot = Heatmap_Strep_A, device = "pdf", width = 20, height = 20)

```